---
title: "Kinetics Overview"
author: "Marc A.T. Teunis, PhD"
date: "`r Sys.time()`"
params:
  path_out: !r here::here('inst', 'final_figures_paper') 
  loose_images: !r here::here('inst', 'final_figures_paper', 'loose_images') 
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{kinetics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE, fig.width = 6, fig.height = 3.5)
```

## Packages
```{r, packages}
library(kinetics)
library(tidyverse)
library(readxl)
# library(fast)
library(Hmisc)
library(car)
library(gtools)
library(pastecs)
library(lme4)
library(nlme)
library(RColorBrewer)
library(agricolae)
library(multcomp)
library(stargazer)
```

## Figure 1
This figure shows the Self-reported Perceived Exertion (RPE) or so-called Borg score over time, during training. The data have been prepared from an Excel file that is included in the './data-raw' folder of the package. It can be loaded as shown below. No exercise is regarded as base-line.

### RPE (Borg scale) scores - data
The RPE data is loaded and cleaned below.
```{r}
data("rpe_scores", package = "kinetics")
sum(is.na(rpe_scores))

map(rpe_scores, unique)

rpe_scores_clean <- rpe_scores %>%
  dplyr::filter(!protocol == "P6")

```

### Colours for the graphs
```{r, colour_palette}
palette <- RColorBrewer::brewer.pal(7, "Set1")
palette
palette_rpe <- c(palette[c(1:3,4)])
```

### RPE plot, first version
```{r, rpe_plot}
sum(is.na(rpe_scores))

# some further transformation
rpe_grouped <- rpe_scores_clean %>%
#  dplyr::filter(!time == 65) %>%
  dplyr::group_by(protocol, time) %>%
  dplyr::summarize(rpe = mean(rpe, na.rm = TRUE)) %>%
  na.omit()

rpe_plot <- ggplot(rpe_grouped,
                   aes(
                     x = time,
                     y = rpe,
                     group = protocol,
                     color = protocol
                   )) +
  geom_line(size = 1.5) +
  expand_limits(y = c(6, 20),
                x = c(-0.4, 6)) +
  labs(x = "Time (minutes)",
       y = "Perceived exertion \n (Borg score)") +
  geom_vline(xintercept = 4,
             linetype = "dashed",
             size = 1) + ## because time is a factor use the ith element on the axis to place the vline +
  
  geom_vline(xintercept = 0,
             linetype = "dashed",
             size = 1) +
  geom_label(aes(x = 0,
                 y = 6,
                 label = "T0"), colour = "black", size = 5) +
  geom_label(aes(x = 4,
                 y = 6,
                 label = "T1"), colour = "black", size = 5) +
  
  scale_color_manual(
    name = "Protocols",
    breaks = c("P2", "P3", "P4", "P5"),
    labels = c("70% Wmax",
               "70% Wmax/DH",
               "50% Wmax",
               "55/85% Wmax"),
    values = palette_rpe
  )  +
  kinetics::theme_individual()

rpe_plot
```

## RPE plot, second version
According rebuttal comments we leave time = 65 min. out of plot and analysis.
```{r, rpe_plot_rebuttal}
rpe_grouped_rebuttal <- rpe_scores_clean %>%
  dplyr::filter(!time == 65) %>%
  dplyr::group_by(protocol, time) %>%
  dplyr::summarize(rpe = mean(rpe, na.rm = TRUE)) %>%
  na.omit()

## new plot - rebuttal
rpe_plot_rebuttal <- ggplot(rpe_grouped_rebuttal,
                            aes(
                              x = time,
                              y = rpe,
                              group = protocol,
                              color = protocol
                            )) +
  geom_line(size = 1.5) +
  expand_limits(y = c(0, 20),
                x = c(-0.4, 5)) +
  labs(x = "Time (minutes)",
       y = "Perceived exertion \n (Borg score)") +
  geom_vline(xintercept = 4,
             linetype = "dashed",
             size = 1) + ## because time is a factor use the ith element on the axis to place the vline +
  
  geom_vline(xintercept = 0,
             linetype = "dashed",
             size = 1) +
  geom_label(aes(x = 0,
                 y = 6,
                 label = "T0"),
             colour = "black", size = 5) +
  geom_label(aes(x = 4,
                 y = 6,
                 label = "T1"), colour = "black", size = 5) +
  
  scale_color_manual(
    name = "Protocols",
    breaks = c("P2", "P3", "P4", "P5"),
    labels = c("70% Wmax",
               "70% Wmax/DH",
               "50% Wmax",
               "55/85% Wmax"),
    values = palette_rpe
  )  +
  kinetics::theme_individual()

rpe_plot_rebuttal

## save plot_rebuttal
ggsave(
  plot = rpe_plot_rebuttal,
  filename = file.path(
    params$path_out,
    paste(Sys.Date(),
          "rpe_plot_rebuttal",
          ".svg",
          sep = "_")
  ),
  width = 11,
  height = 7,
  #  dpi = 300,
  #  units = "cm"
)

ggsave(
  plot = rpe_plot_rebuttal,
   filename = file.path(
    params$path_out,
    paste(Sys.Date(),
          "rpe_plot_rebuttal",
          ".png",
          sep = "_")
  ),
  width = 11,
  height = 7,
  #  dpi = 300,
  #  units = "cm"
)

ggsave(
  plot = rpe_plot_rebuttal,
   filename = file.path(
    params$path_out,
    paste(Sys.Date(),
          "rpe_plot_rebuttal",
          ".jpeg",
          sep = "_")
  ),
  width = 11,
  height = 7,
  #  dpi = 300,
  #  units = "cm"
)

rpe_plot_rebuttal
```

## Statistical analysis of RPE

### Assumptions & Diagnostics
```{r}

## normality data
rpe_scores_clean %>%
  ggplot(
    aes(
      x = rpe)) +
#  geom_histogram() +
  geom_density()
    
rpe_scores_clean %>%
  ggplot(
    aes(
      x = rpe)) +
  geom_histogram() 
#  geom_density()

## shapiro - wilk
shapiro.test(rpe_scores$rpe)
## transforming makes it worse
shapiro.test(log(rpe_scores$rpe))

## We will investigate normality of residuals furtheron

## diagnostic plots
## individual observations


rpe_scores_clean %>%
  dplyr::filter(!time == 65) %>%
  ggplot(aes(
    x = time,
    y = rpe)) +
  geom_point(aes(colour = protocol), position = "jitter") +
  geom_hline(
    data = rpe_grouped_rebuttal %>% dplyr::filter(!time == 65), 
    aes(
      yintercept = rpe,
      colour = protocol), linetype = "dashed", size = 1) +
  ggrepel::geom_label_repel(aes(label = subject), size = 2.5) +
  facet_wrap(~time, ncol = 4)


```

### Inference
```{r}

rpe_scores_rebuttal <- rpe_scores %>%
  dplyr::filter(!time == 65,
                !protocol == "P6") %>%
  droplevels()

map(rpe_scores_rebuttal, levels)


## mixed effects model
 model <-  lme(rpe ~ time * protocol, 
                    data = rpe_scores_rebuttal, 
                    method = "ML", 
                    na.action = "na.omit", 
                    random = ~1 | subject)

  # model_summary <- summary(model) 
  anova(model)
  summary(model)

  ## post hoc
  summary(glht(model, linfct=mcp(time="Tukey")))
  summary(glht(model, linfct=mcp(protocol="Tukey")))
  
  ## investigate interactions and co-variates
#  rpe_scores_rebuttal$protocol_time_interaction <- interaction(
#    rpe_scores_rebuttal$protocol, rpe_scores_rebuttal$time)
#model <- lme(rpe ~ protocol_time_interaction, 
#             random = ~1|subject/protocol/time, data=rpe_scores_rebuttal %>% na.omit()) 
#comp.timegeno <- glht(model, linfct=mcp(protocol_time_interaction="Tukey")) 
#summary(comp.timegeno)  
  
  
plot(model)
  

```

## Does the conclusion change if we use a hierachical model with time nested in protocol nested in subject
```{r}
## mixed effects model
 model2 <-  lme(rpe ~ time * protocol, 
                    data = rpe_scores_rebuttal, 
                    method = "ML", 
                    na.action = "na.omit", 
                    random = ~1 | subject/protocol/time) ## add nested structure /pro/tim

  # model_summary <- summary(model) 
  anova(model2)
  summary(model2)

  
  plot(model2)
  
plot(model2$fitted, model2$residuals)  
  
  
## compare models  
anova(model, model2)  

  ## post hoc
 
  summary(glht(model, linfct=mcp(time="Tukey")))
  summary(glht(model, linfct=mcp(protocol="Tukey")))

 hist(model2$residuals)
 shapiro.test(model2$residuals)
  
# model_summary <- summary(model) 
  anova(model2)
  summary(model2)

anova(model, model2)
#  plot(model2)
  
plot(model2$fitted, model2$residuals)  
  
  
  ## post hoc
 
  summary(glht(model2, linfct=mcp(time="Tukey")))
  summary(glht(model2, linfct=mcp(protocol="Tukey")))

  

  
```

## Model 3 - simpeler (linear model, unpaired / disregards the paired nature of the experiment)
```{r}
## lm model
model3 <- lm(rpe ~ time * protocol, data = rpe_scores_rebuttal) 
anova(model3)
summary(glht(model3, linfct=mcp(protocol="Tukey")))

  
plot(model3)
  
plot(model3$fitted, model3$residuals)  
  
  
## compare models  
x <- anova(model, model2, model3)
x

  ## post hoc

  summary(glht(model, linfct=mcp(time="Tukey")))
  summary(glht(model, linfct=mcp(protocol="Tukey")))
```

## Model 4 - as model 1, but log transformed RPE values
```{r}
## lm model
 model4 <-  lme(log(rpe) ~ time * protocol, 
                    data = rpe_scores_rebuttal, 
                    method = "ML", 
                    na.action = "na.omit", 
                    random = ~1 | subject)

anova(model4)
summary(glht(model4, linfct=mcp(protocol="Tukey")))

  
#plot(model4)
#plot(model)
  
hist(model4$residuals)

  ## post hoc
 
  summary(glht(model, linfct=mcp(time="Tukey")))
  summary(glht(model, linfct=mcp(protocol="Tukey")))
```


## Model 5 - as model 4, but log transformed RPE values and nested
```{r}
## lm model
 model5 <-  lme(log(rpe) ~ time * protocol, 
                    data = rpe_scores_rebuttal, 
                    method = "ML", 
                    na.action = "na.omit", 
                    random = ~1 | subject/protocol/time)

anova(model5)
summary(glht(model5, linfct=mcp(protocol="Tukey")))

  
#plot(model5)
#plot(model)
  
hist(model4$residuals)


  summary(glht(model, linfct=mcp(time="Tukey")))
  summary(glht(model, linfct=mcp(protocol="Tukey")))
  
anova(model4, model5)  

qqnorm(model5$residuals)  
qqline(model5$residuals)
qqnorm(model4$residuals)  
qqline(model4$residuals)
qqnorm(model$residuals)
qqline(model$residuals)


```

## CONCLUSIONS:

The model nummer 1 is the best model (the 'simple' mixed effects model) without nesting of subject in protocol in time, where rpe is not log transformed. The anova between the models reveals this as well as the residuals plots and the qqplot of the residuals of model1

```
Model - Fit: 
lme.formula(
  fixed = rpe ~ time * protocol, 
  data = rpe_scores_rebuttal, 
  random = ~1 | subject, 
  method = "ML", na.action = "na.omit"
  )
```



The residual plot of model1 shows the least systematic structure
```{r}
plot(model$residuals, model$fitted)
```

Normality check on the residuals reveals that there is a significant P, 
which suggests deviation from normality.

H0 for Shapiro -> observations were drawn from a Normal distribution vs
H.alt observation were not drawn from a normal distribution.
```{r}
shapiro.test(model$residuals)
qqnorm(model$residuals)
qqline(model$residuals)
hist(model$residuals) ## the distribution of the residuals is somewhat skewed
```

The models main effects
```{r}
anova(model)
summary(model)
```

The models pair-wise comparisons
see: https://stats.stackexchange.com/questions/5250/multiple-comparisons-on-a-mixed-effects-model
```{r}
summary(glht(model, linfct=mcp(time="Tukey")))
summary(glht(model, linfct=mcp(protocol="Tukey")))


```

## Write out statitics to summary file
```{r}
model_main <- anova(model)
model_contrasts <- list(
  summary(glht(model, linfct=mcp(time="Tukey"))),
  summary(glht(model, linfct=mcp(protocol="Tukey"))))


stargazer::stargazer(model_main)

write_lines(model_main %>% broom::tidy(), path = 
                 file.path(
                 params$path_out,
                 "statistics_rpe.txt"))

```

## Heatmaps from model metrics

### Statistical metrics data
Effects sizes and probabilities (F values and P values) were calculated using a 
multi-level mixed effects model approach as reported earlier ([DOI:(https://doi.org/10.1016/j.clnu.2018.09.029)])
Here we load the F- and P- values for the analytes discussed in the kinetics 
paper 

#### F-values data
```{r}
data("f_values_long_protocol", package = "kinetics")
f_values_long_protocol
```

#### P-values data
```{r}
data("p_values_long_protocol", package = "kinetics")
p_values_long_protocol
```


### Inspect data
F- values
```{r, inspect_statistics_data}
names(f_values_long_protocol)

levels(f_values_long_protocol$analyte)

f_values_long_protocol$contrast <- as_factor(f_values_long_protocol$contrast)

levels(f_values_long_protocol$contrast)

f_values_final <- f_values_long_protocol %>%
  mutate(statistic = "f_value") %>%
  print()
```

P-values
```{r}
names(p_values_long_protocol)
levels(p_values_long_protocol$analyte)

p_values_long_protocol$contrast <- as_factor(p_values_long_protocol$contrast)
levels(p_values_long_protocol$contrast)

p_values_final <- p_values_long_protocol

ind <- names(f_values_final) == names(p_values_final)
all <- all(ind)
if(all == "FALSE"){
  stop("ERROR, names of dataframes do not match, cannot bind rows - FIX before proceeding")
}

```


Combine F- and P-values
```{r}
## bind rows
f_and_p <- dplyr::bind_rows(f_values_final,
                            p_values_final) %>%
  print()

levels(as_factor(f_and_p$statistic))

analytes_f <- f_values_final$analyte %>%
  as_factor() %>%
  levels()

## combine
both <- f_and_p %>%
  dplyr::filter(analyte %in% analytes_f)


analytes_final <- as_factor(both$analyte) %>% levels()
analytes_final

# usethis::use_data(both)
```

## Heatmap graps

```{r, individual_heatmaps, include=FALSE}
heatmap_p_values <- both %>%
  dplyr::filter(statistic == "p_value") %>%
  mutate(log10_p = log10(value)) %>%
  arrange(desc(value)) %>%
  
#  as_tibble() %>%
ggplot(., aes(x = contrast, 
              y = reorder(analyte, desc(value)))) +
  geom_tile(aes(fill = log10_p), color = "white") +
  scale_fill_gradient(low = "darkgreen", high = "white") +
  
  ylab("Analyte") +
  xlab("Contrast") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
   theme(plot.margin=unit(c(2,2,2,2),"cm"))
heatmap_p_values

# scale_y_discrete(limits = rev(levels())) +
#  scale_x_discrete()[order(.$analyte_new_lbl)] +
#  labs(fill = "F value") 
#facet_wrap(~ statistic)

# heatmap_p_values

 ggsave(plot = heatmap_p_values,  filename = file.path(
    params$path_out,
    paste(Sys.Date(),
    "heatmap_p_values_protocol",
    ".svg", 
    sep = "_" )
    ), 
        width = 18, 
        height = 18, 
        dpi = 300, 
        units = "cm"
        )
 
 ggsave(plot = heatmap_p_values,
        filename = file.path(
    params$path_out,
                             paste(Sys.Date(),
                                   "heatmap_p_values_protocol",
                                   ".png", 
                                   sep = "_" )
                             ), 
        width = 18, 
        height = 18, 
        dpi = 300, 
        units = "cm"
        )

heatmap_f_values <- both %>%
  dplyr::filter(statistic == "f_value") %>%
  arrange(desc(value)) %>%
  
#  as_tibble() %>%
ggplot(., aes(x = contrast, 
              y = reorder(analyte, value))) +
  geom_tile(aes(fill = value), color = "white") +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  
  ylab("Analyte") +
  xlab("Contrast") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
   theme(plot.margin=unit(c(2,2,2,2),"cm"))
heatmap_f_values

 ggsave(plot = heatmap_f_values,
        filename = file.path(
    params$path_out,
                             paste(Sys.Date(),
                                   "heatmap_f_values_protocol",
                                   ".svg", 
                                   sep = "_" )
                             ), 
        width = 14, 
        height = 18, 
        dpi = 300, 
        units = "cm"
        )
 
 ggsave(plot = heatmap_f_values,
        filename = file.path(
    params$path_out,
                             paste(Sys.Date(),
                                   "heatmap_f_values_protocol",
                                   ".png", 
                                   sep = "_" )
                             ), 
        width = 18, 
        height = 18, 
        dpi = 300, 
        units = "cm"
        )

```

## Figure 2
This is the code for the heatmap in Figure 2 of the manuscript
```{r, fig.width=12, fig.height=24}

invalid_params <- c("gfra", "ip10", "facd", "crph", "tnfa")


## F-values
heatmap_f_values <- both %>%
  dplyr::filter(statistic == "f_value") %>%
  dplyr::filter(!analyte %in% invalid_params) %>%
  arrange(desc(value))

heatmap_f_values
## P-values  

## function to convert p-value to stars
signif_num <- function(x) {
    symnum(x, corr = FALSE, na = FALSE, legend = FALSE,
           cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), 
           symbols = c("***", "**", "*", "#", "+"))
}

heatmap_p_values <- both %>%
  dplyr::filter(statistic == "p_value") %>%
  dplyr::filter(!analyte %in% invalid_params) %>%
  mutate(
    original_value = (value/1000),
    log10_p = log10(value),
         stars_noquote = signif_num(original_value)) %>%
  mutate(stars = as.character(stars_noquote)) %>%
    arrange(desc(value))

#vheatmap_p_values

#heatmap_p_values %>%
#  dplyr::filter(analyte == "lys")


# checking other duplicates
# names(heatmap_p_values)
x <- heatmap_p_values %>%
  dplyr::select(analyte, contrast) %>%
  group_by(analyte) %>%
  nest()

how_many_contrasts <- function(i){
i %>% nrow()
}

x <- x %>%
  mutate(number_contrasts = map(data, how_many_contrasts))

# x$number_contrasts

#nrow(x)
#unique(x) %>% nrow()

#  as_tibble() %>%
with_stars <- ggplot(data = heatmap_f_values, aes(x = contrast, 
              y = reorder(analyte, value))) +
  geom_tile(aes(fill = value), color = "white") +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  geom_text(data = heatmap_p_values,
            aes(label = sprintf("%1.3s", stars)), vjust = 0.6, size = 3) +
  ylab("Parameter") +
  xlab("Comparison") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
   theme(plot.margin=unit(c(2,2,2,2),"cm")) +
  theme_bw() +
  labs(fill = "F Value")

with_stars
## save to disk

# ggsave(plot = with_stars,
#         filename = file.path(
#     params$path_out,
#                              paste(Sys.Date(),
#                                    "heatmap_f_values_protocol_with_stars",
#                                    ".svg", 
#                                    sep = "_" )
#                              ), 
#         width = 24, 
#         height = 18, 
#         dpi = 600, 
#         units = "cm"
#         )

# ggsave(plot = with_stars,
#         filename = file.path(
#     params$path_out,
#                              paste(Sys.Date(),
#                                    "heatmap_f_values_protocol_with_stars",
#                                    ".png", 
#                                    sep = "_" )
#                              ), 
#         width = 24, 
#         height = 18, 
#         dpi = 600, 
#         units = "cm"
#         )

# ggsave(plot = with_stars,
#         filename = file.path(
#     params$path_out,
#                              paste(Sys.Date(),
#                                    "heatmap_f_values_protocol_with_stars",
#                                    ".jpeg", 
#                                    sep = "_" )
#                              ), 
#         width = 24, 
#         height = 18, 
#         dpi = 600, 
#         units = "cm"
#         )


# heatmap_p_values$stars_noquote %>% unique()




```

## Some data checking and transformations
```{r, eval=FALSE}
histograms <- both %>%
 ggplot(aes(value)) + 
  geom_density(aes(y=..density..), 
                 colour="black", 
                 fill="white") + 
  labs(x="Value", y = "Density") +
  theme(legend.position = "none") +
  facet_wrap(~statistic)
histograms
```

```{r, eval=FALSE}
histograms <- both %>%
 ggplot(aes(log10(value))) + 
  geom_density(aes(y=..density..), 
                 colour="black", 
                 fill="white") + 
  labs(x="Value", y = "Density") +
  theme(legend.position = "none") +
  facet_wrap(~statistic)
histograms
```

```{r, eval=FALSE}
histograms <- both %>%
 ggplot(aes(sqrt(value))) + 
  geom_density(aes(y=..density..), 
                 colour="black", 
                 fill="white") + 
  labs(x="Value", y = "Density") +
  theme(legend.position = "none") +
  facet_wrap(~statistic)
histograms
```

```{r, eval=FALSE}
histograms <- both %>%
 ggplot(aes(value^2)) + 
  geom_density(aes(y=..density..), 
                 colour="black", 
                 fill="white") + 
  labs(x="Value", y = "Density") +
  theme(legend.position = "none") +
  facet_wrap(~statistic)
histograms
```


```{r, eval=FALSE}
histograms <- both %>%
 ggplot(aes(1/value)) + 
  geom_density(aes(y=..density..), 
                 colour="black", 
                 fill="white") + 
  labs(x="Value", y = "Density") +
  theme(legend.position = "none") +
  facet_wrap(~statistic)
histograms
```


```{r, eval=FALSE}

heatmap_f_values <- both %>%
  dplyr::filter(statistic == "f_value") %>%
  mutate(sqrt_f = sqrt(value)) %>%
  arrange(desc(value)) %>%
  
#  as_tibble() %>%
ggplot(., aes(x = contrast, 
              y = reorder(analyte, value))) +
  geom_tile(aes(fill = sqrt_f), color = "white") +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  
  ylab("Analyte") +
  xlab("Contrast") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
   theme(plot.margin=unit(c(2,2,2,2),"cm"))
heatmap_f_values


heatmap_p_values <- both %>%
  dplyr::filter(statistic == "p_value") %>%
  mutate(sqrt_p = sqrt(value)) %>%
  arrange(desc(value)) %>%
  
#  as_tibble() %>%
ggplot(., aes(x = contrast, 
              y = reorder(analyte, desc(value)))) +
  geom_tile(aes(fill = sqrt_p), color = "white") +
  scale_fill_gradient(low = "darkgreen", high = "white") +
  
  ylab("Analyte") +
  xlab("Contrast") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
   theme(plot.margin=unit(c(2,2,2,2),"cm"))
heatmap_p_values

# scale_y_discrete(limits = rev(levels())) +
#  scale_x_discrete()[order(.$analyte_new_lbl)] +
#  labs(fill = "F value") 
#facet_wrap(~ statistic)

heatmap_p_values

panel <- cowplot::plot_grid(heatmap_p_values,
                                         heatmap_f_values,
                            labels = c("A", "B"), vjust = 4, hjust = -8
                            )

panel
 ggsave(plot = panel,
        filename = here::here(
                             "inst",
                             "images",
                             paste(Sys.Date(),
                                   "heatmap_panel_sqrt",
                                   ".svg", 
                                   sep = "_" )
                             ), 
        width = 40, 
        height = 30, 
        dpi = 300, 
        units = "cm"
        )

ggsave(plot = panel,
        filename = here::here(
                             "inst",
                             "images",
                             paste(Sys.Date(),
                                   "heatmap_panel_sqrt",
                                   ".png", 
                                   sep = "_" )
                             ), 
        width = 40, 
        height = 30, 
        dpi = 300, 
        units = "cm"
        )

ggsave(plot = panel,
        filename = here::here(
                             "inst",
                             "images",
                             paste(Sys.Date(),
                                   "heatmap_panel_sqrt",
                                   ".jpeg", 
                                   sep = "_" )
                             ), 
        width = 40, 
        height = 30, 
        dpi = 300, 
        units = "cm"
        )

```

## Prepare data for analyte kinetics graphs

### Load data for analytes
```{r, panel_lines}

library(kinetics)

data("analyte_data", package = "kinetics")

data_summary <- analyte_data %>%
  arrange(protocol, analyte)
data_summary


data_summary$analyte <- as.factor(data_summary$analyte)
levels(data_summary$analyte)

data_split <- split(data_summary, 
                        droplevels(data_summary$analyte))




```

### Load analyte annotations
```{r}
data("analyte_annotations", package = "kinetics")
```

## Colours
```{r}
palette <- RColorBrewer::brewer.pal(7, "Set1")
palette <- c(palette[c(1:3,4,5)])
```


## Line plots
```{r, all_lines, eval=TRUE}

image_directory <- params$loose_images
dir.create(image_directory)

# argument for function test
analyte = "ala"
kinetics::draw_lines(DF = data_split[[1]], palette_graph = palette)

test_lines <- print_lines(analyte = "SODIUM")
#test_lines


test_lines <- print_lines(analyte = "INSULINE")
#test_lines

run_plotlist <- function(){
  plot_list <- lapply(levels(data_summary$analyte), print_lines)
  names(plot_list) <- levels(data_summary$analyte)
  names(plot_list)
}


```

## Panels
```{r, panels_plot_list, eval=FALSE}

## test
kinetics::print_lines_panel("GLU_NS", x_pos = c(1,2))
print_lines_panel("POTASSIUM", x_pos = c(1,2))

```

## Prepare potlist for panels
```{r, plotlist}
## generate plotlist
analytes <- c(
  "BICARB",
  "KREAT",
  "POTASSIUM",
  "SODIUM",
  "CHLORIDE",
  "ALB",
  "INSULINE",
  "GLU_NS",
  "LEUCO",
  "ERY",
  "HB",
  "HT")

analyte_data <- data_summary %>%
  dplyr::filter(analyte %in% analytes) %>%
  droplevels()

levels(analyte_data$analyte)



x_pos_list <- list(
  "alb" = c(1,3),
  "bicarb" = c(1,3),
  "chloride" = c(1,3),
  "ery" = c(1,3),
  "glu_ns" = c(1,2), ## the GLU_NS set has less time points
  "hb" = c(1,3),
  "ht" = c(1,3),
  "insuline" = c(1,3), 
  "kreat" = c(1,3),
  "leuco" = c(1,3),
  "potassium" = c(1,3),
  "sodium" = c(1,3)
)
x_pos_list



## select analytes relevant for kinetics paper - panels
analyte_data <- data_summary %>%
  dplyr::filter(analyte %in% analytes) %>%
  droplevels()

levels(analyte_data$analyte)

panel_plot_list <- map2(
  .x = levels(analyte_data$analyte), 
  .y = x_pos_list,
  .f = print_lines_panel
  )

# panel_plot_list
## check x_pos for GLU_NS
names(panel_plot_list) <- levels(analyte_data$analyte)
names(panel_plot_list)

## save to disk for use in the package
# usethis::use_data(panel_plot_list)

```

## Final panels
```{r}
## Panels

## get panel plotlist

#panel_plot_list <- readr::read_rds(
#  path = here::here("data", "panel_plot_list.rds")
#  )

#•	Bicarbonate, kreatinine, potassium, sodium, chloride
#•	Albumine, insulin, glucose 
#•	Leukocytes, erythrocytes, Hb, Ht
## read panel_plot_list from disk

#panel_plot_list <- readr::read_rds(path = here::here(
#  "data",
#  "panel_plot_list.rds"
#  ))

# figure 1; panel
p1_a <- panel_plot_list[["BICARB"]]
#p1_a
p1_b <- panel_plot_list[["KREAT"]]
p1_c <- panel_plot_list[["POTASSIUM"]]
p1_e <- panel_plot_list[["SODIUM"]]
p1_d <- panel_plot_list[["CHLORIDE"]]

# figure 2; panel
p2_a <- panel_plot_list[["ALB"]]
p2_b <- panel_plot_list[["INSULINE"]]
p2_c <- panel_plot_list[["GLU_NS"]]

# figure 3; panel
p3_a <- panel_plot_list[["LEUCO"]] + ylab("Counts")
p3_b <- panel_plot_list[["ERY"]] + ylab("Counts")
p3_c <- panel_plot_list[["HB"]]
p3_d <- panel_plot_list[["HT"]]


# figure 4; panel
#p4_a <- panel_plot_list[["CHLORIDE"]]
#p4_b <- panel_plot_list[["ALB"]]
#p4_c <- panel_plot_list[["ERY"]]

## Panels

## dimensions
height <- 26
width <- 28
resolution <- 600
```

## Figure 3
```{r, fig_3, fig.width=width, fig.height=height/2, dpi=resolution}
###Figure 1

library(cowplot)
library(svglite)
# https://cran.r-project.org/web/packages/cowplot/vignettes/shared_legends.html

horizontal_label_offset <- -7
vertical_label_offset <- 4
label_size <- 20

figure_1 <- cowplot::plot_grid( p1_a + theme(legend.position="none"),
           p1_b + theme(legend.position="none"),
           p1_c + theme(legend.position="none"),
           p1_d + theme(legend.position="none"),
           p1_e + theme(legend.position="none"),
           align = 'vh',
       #    labels = c("A", "B", "C", "D", "E"),
           hjust = horizontal_label_offset, 
           vjust = vertical_label_offset,
           ncol = 3,
           label_size = label_size
           )

#cfigure_1


legend_1 <- get_legend(p1_a + theme(legend.position="bottom"))

p_1 <- plot_grid(legend_1, figure_1, ncol = 1, rel_heights = c(0.05, 1))
p_1
## citrulliner::save_in_image_directory(filename = "test2.svg", height = 26, width = 7)
```

```{r}
ggsave(p_1, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "kinetics_fig_1.svg", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)

ggsave(p_1, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "kinetics_fig_1.png", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)


ggsave(p_1, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "kinetics_fig_1.jpeg", sep = "_")), 
       height = height/2.2, 
       width = width/1.5, 
       dpi = resolution/2)


```

## Figure 4
```{r, fig_4, fig.width=width, fig.height=height/2, dpi=resolution}
### Figure 2

figure_2 <- plot_grid( p2_a + theme(legend.position="none"),
           p2_b + theme(legend.position="none"),
           p2_c + theme(legend.position="none"),
           align = 'vh',
      #     labels = c("A", "B", "C"),
           hjust = horizontal_label_offset,
           vjust = vertical_label_offset-1,
           ncol = 3,
           label_size = label_size
           )

# figure_2


legend_2 <- get_legend(p2_a + theme(legend.position="bottom"))

p_2 <- plot_grid(legend_2, figure_2, ncol = 1, rel_heights = c(0.05, 1))

p_2
## citrulliner::save_in_image_directory(filename = "test2.svg", height = 26, width = 7)
```

```{r}
ggsave(p_2, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "kinetics_fig_2.svg", sep = "_")), 
       height = height/4.5, 
       width = width/1.5,
       dpi = resolution/2)

ggsave(p_2, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "kinetics_fig_2.png", sep = "_")), 
       height = height/4.5, 
       width = width/1.5,
       dpi = resolution/2)

ggsave(p_2, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "kinetics_fig_2.jpeg", sep = "_")), 
       height = height/4.5, 
       width = width/1.5,
       dpi = resolution/2)


```

### Figure 5
```{r, fig_5, fig.width=width/2, fig.height=height, dpi=resolution}
figure_3 <- plot_grid( p3_a + theme(legend.position="none"),
           p3_b + theme(legend.position="none"),
           p3_c + theme(legend.position="none"),
           p3_d + theme(legend.position="none"),
           align = 'vh',
        #   labels = c("A", "B", "C", "D"),
           hjust = horizontal_label_offset,
           vjust = vertical_label_offset,
           ncol = 2,
           label_size = label_size
           )

# figure_3


legend_3 <- get_legend(p3_a + theme(legend.position="bottom"))

p_3 <- plot_grid(legend_3, figure_3, ncol = 1, rel_heights = c(0.05, 1))

## citrulliner::save_in_image_directory(filename = "test2.svg", height = 26, width = 7)
p_3
```

```{r}
ggsave(p_3, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "kinetics_fig_3.svg", sep = "_")), 
       height = height/2.4,
       width = width/2.4,
       dpi = resolution/2)

ggsave(p_3, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "kinetics_fig_3.png", sep = "_")), 
       height = height/2.4,
       width = width/2.4,
       dpi = resolution/2)


ggsave(p_3, filename = file.path(
    params$path_out,
                         paste(Sys.Date(),
                         "kinetics_fig_3.jpeg", sep = "_")), 
       height = height/2.4,
       width = width/2.4,
       dpi = resolution/2)




```

## Summary table
```{r}

## all analytes in data
unique(data_summary$analyte) %>%
  as.character %>%
  sort %>%
  print()

analytes_heatmap <- unique(heatmap_f_values$analyte) %>%
  as.character() %>%
  sort() %>%
  print()

analytes_paper <- c(
  
  # line figures
  "BICARB", #1
  
  "KREAT", #2
  "POTASSIUM", #3
  "SODIUM", #4
  "CHLORIDE", #5
  
  "ALB", #6
  "INSULINE", #7
  "GLU_NS", #8
  
  "LEUCO", #9
  "ERY", #10
  "HB", #11
  "HT", #12

  ## heatmap  
  "ala",
  "arg",
  "asp",
  "asn",
  "citrul",
  "CORT",
  "glu",
  "gln",
  "gly",
  "ifabp",
  "ile",
  "leu",
  "LYMFO_ABS_CORR",
  "lys",
  "meth",
  "MONO_ABS",
  "NEUTRO_ABS",
  "pheala",
  "ser",
  "tau",
  "thre",
  "trp",
  "tyr",
  "UREUM",
  "val"   
)

analytes_paper %>% sort

length(analytes_paper)
  
data_paper_summary <- data_summary %>%
  dplyr::filter(analyte %in% analytes_paper) %>%
  print()

unique(data_paper_summary$analyte) %>% as.character() %>% sort %>% length()  
analytes_paper %>% sort %>% length()

analytes_paper %in% unique(data_paper_summary$analyte)

## add rpe
rpe_summary <- rpe_scores_clean %>%
  group_by(protocol, time) %>%
 summarise(mean_conc = round(mean(rpe, na.rm = TRUE), digits = 3),
            sd = round(sd(rpe, na.rm = TRUE), digits = 2),
            n_obs = n()) %>%
  dplyr::filter(time != 65) %>%
  dplyr::rename(measured = mean_conc) %>%
  dplyr::mutate(analyte = "RPE") %>%
  dplyr::select(protocol, time, analyte, measured, sd, n_obs)

rpe_summary  

## REname
data_paper_summary <- data_paper_summary %>%
  dplyr::rename(measured = mean_conc)

## combine RPE with rest
data_paper_summary_all <- 
  dplyr::bind_rows(
    rpe_summary,
    data_paper_summary
  ) %>% 
  dplyr::rename(mean = measured) %>%
  arrange(protocol, analyte, time)







data_paper_summary_all$analyte %>% unique()

## add zonuline
source(
  here::here(
    "R",
    "data_clean_grinta_zonulin.R"
  )
)

zonulin <- data_clean_grinta_zonulin()

## clean op for join
names(data_paper_summary_all)


zonulin_clean <- zonulin %>%
  dplyr::mutate(time = as.character(time)) %>%
  dplyr::rename(analyte = analyse) %>%
  dplyr::select(subject,
         protocol,
         time,
         analyte,
         result)
zonulin_clean

zonulin_summary <- zonulin_clean %>%
  group_by(protocol, time, analyte) %>%
  summarise(mean = round(mean(result, na.rm = TRUE), digits = 3),
            sd = round(sd(result, na.rm = TRUE), digits = 2),
            n_obs = n())


zonulin_summary

names(zonulin_summary) == names(data_paper_summary_all)

## row bind zonuline to summary data
data_paper_all_final <- dplyr::bind_rows(
  data_paper_summary_all,
  zonulin_summary
) %>%
arrange(analyte, protocol, time)  

data_paper_all_final

# usethis::use_data(data_paper_all_final)

# save to disk
# install.packages("xlsx", dependencies = TRUE)
readr::write_csv(data_paper_all_final, 
                 path = file.path(
                   params$path_out,
                 "summary_data_kinetics.csv"
                 )
                 )
                
```
